Circumpolar variations in the chaotic nature of Southern Ocean eddy dynamics

### 2022-04-17 revised_paper/ResponseToReviewers.tex

\documentclass[11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper,
  top=25mm,
  bottom=30mm,
  left=25mm,
  right=25mm}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{color}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

@@

\title{Brief Article}
\author{The Author}
%\date{}                                           % Activate to display a given date or no date

\begin{document}
%\maketitle

\noindent We would like to convey our thanks to both reviewers for the constructive and positive reviews of our paper. 
In the following we have addressed each of the reviewers' comments -- with their comments in blue and our response in plain text.

@@

Overall, the question is important, the approach is creative, the findings are significant, and the presentation is high quality. I think the paper is certainly publishable in JGR. I only have a few minor comments and questions, which I hope the authors can clarify before my recommendation of acceptance.\\}

\noindent Our thanks to the reviewer for their positive assessment. 
We have responded in detail to each of the reviewer's comments below.

\subsection*{\color{blue}Details:}

{\color{blue} 
\begin{description}
\item[ Line 84: ] ``the Southern Ocean", missing words like EKE or eddy field?\\
{\color{black}  We agree and have changed to read ``Southern Ocean EKE''.}

@@

\item[ Lines 140-141:] ``the member's global mean sea level anomaly", is this a single value or a time series? If this is a value, it's more likely to shift each ensemble member's sea surface height with a different value. But it wouldn't affect the sea level anomaly time series at each grid point. \\
{\color{black} This statement refers to a timeseries of the globally averaged sea level anomaly. We have clarified this statement to read:\\
 ``For each ensemble member, and at every time step, the member's global mean sea level anomaly is subtracted at every grid point  \ldots" }

\item[Lines 158-162:] The authors mention that the seasonal EKE variation dominates the statistics. To look at the forced response or the correlation between ensemble-mean EKE and wind stress, they deseasonalize the EKE time series. There is some missing information about how they smooth wind stress, the same as the deseasonalized EKE? Also, maybe some general information about the wind stress can be provided in the method section.\\
{\color{black} Yes, the seasonal cycle is also removed from the wind stress. We have added, on line 189, the caveat that wind stress ``\ldots has been deseasonalised similarly to $E_i$ \ldots". }

@@

\item[Lines 176-179:] ``Analogously, the intrinsic variance emerges from the variance, indicating the spread of each member from the ensemble mean"; I feel this sentence is not accurate. Normally the closest meaning of spread is the standard deviation. I would suggest the author rephrase it like this: ``For each time step, the intrinsic variance emerges from the variance of EKE difference between ensemble mean and each member of the 50 simulations.'' \\
{\color{black} Amended as suggested. }

\item[Lines 185-186:] between the EKE (...) AND WHAT? Missing words?\\
{\color{black} We have added ``and its forcing''. }

\item[Line 201:] show $\to$ shows\\
{\color{black} Amended. }

\item[Figure 1:] The ensemble mean EKE pattern is consistent with the mean EKE pattern from satellite altimetry. Since this study focuses more on the EKE variabilities, it would be helpful showing to what extent the averaged EKE anomaly time series (group members and ensemble mean) agree with the EKE anomaly time series from satellite altimetry. Also, please label the longitude, latitude, and the major topographic features. \\
{\color{black}  We have added latitude and longitude to Figure 1 as suggested (and made minor improvements to the plot).
We tested labelling major topographic features, but there are many features that could be labelled and it quickly gets messy, so opted to leave the plot as is.

Including the observed EKE trend is a good idea, and we have added this data to Figure~2(a) (green line), using EKE data calculated from Mart\`inez-Moreno et al. (2021).
This new plot shows that the observed EKE is indistinguishable from the individual ensemble members, but differs from the ensemble mean.}

@@

\item[Line 231:] add citations (e.g., Meredith \& Hogg, 2006; Hogg et al., 2015) \\
{\color{black} Added as suggested }

\item[Line 231:] As found in this paper, the significance of the wind-leading lag correlation is affected by the size and location of the research area. To compare with previous studies, such as Meredith \& Hogg (2006), Hogg et al., (2015), doesn't it make more sense to use the same average boxes? \\
{\color{black} This line actually refers to the circumpolar average, and so provides a direct comparison, as the reviewer suggests.  
For the later parts of the paper where we examine sub-domains, we considered replicating the original Meredith \& Hogg (2006) boxes, but we found that the ensemble approach allows us to refine those boxes to be considerably smaller, while still obtaining meaningful results.
We could happily add an extra figure with the larger Meredith \& Hogg (2006) boxes, but considered this to be of marginal value to the reader.}

\item[Figure 2:] Based on the definition of EKE and wind stress, they should always be positive. To be accurate, I would suggest the authors label them clearly. Same comment for time series plots in figure 3 and figure 4. \\
{\color{black} We agree more information is needed here and have noted in the subtitle of figure 2(a) and the caption of figure 2 that the timeseries shows ``ensemble mean {\color{red}EKE anomaly} (red) along with wind stress {\color{red}anomaly} \ldots'', where red text has been added. 
The captions for figures 3 and 4 have been similarly clarified.}

\item[Line 239-240:] What's the T value for the 4 months and 30 months peaks? \\
{\color{black} The T-values in this case are 5.9 and 3.5 respectively. 
These values have been added to the manuscript as suggested.  }

@@

\item[Line 241:] dominanent $\to$ dominant?\\
{\color{black} Corrected, thankyou. }

\item[Line 274:] Southeast $\to$ Southwest\\
{\color{black} Amended. }

\item[Figure 3:] (a) A colorbar could be useful. Also, when you define the 3 boxes, do you use rectangular boxes like (70$^\circ$E-120$^\circ$E, 45$^\circ$S-60$^\circ$S)? The boxes shown in Figure 3(a) are more like irregular polygons. It will be better if the author can describe how those boxes are defined. \\
{\color{black}  We have added a colorbar as suggested. Regarding the irregular polygons -- yes, the boxes are long lines of latitude, but we were a little lazy in plotting these on the map projection. These boxes have now been fixed to follow latitude lines. }

### 2022-04-17 revised_paper/revised.tex
tinkering with text and responses

The surface eddy kinetic energy for each ensemble member is calculated as 
\begin{linenomath*}
\begin{equation}
E_i = \frac{1}{2}(u_i'^2 + v_i'^2),
\end{equation}
\end{linenomath*}
and subsampled at monthly temporal resolution through weighted averages.
It is notable that $E_i$ contains a large component of seasonal variation which dominates the statistics \citep{Martinez-Moreno2022}.
To look at the forced response, we therefore deseasonalise $E_i$, by removing the mean seasonal cycle.
We then proceed to compute ensemble statistics from this deseasonalised EKE.

@@

To examine the relationship between the EKE (of either the ensemble mean, or from individual ensemble members) and its forcing, we are guided by previous work which emphasises the role of wind stress in governing the eddy variability at different timescales \citep[e.g.][]{Hogg2015, Sinha2016}.
Thus, we look at the correlation coefficient, $r$, at a range of lags between the wind stress $|\tau|$ (which is identical for all ensemble members and has been deseasonalised in a similar way to $E_i$) and EKE.
For each correlation we evaluate the statistical significance of the correlation \citep[following, e.g.][]{Santer2000}.
We first calculate the effective sample size, $N_e$, where
\begin{linenomath*}
\begin{equation}
N_e \equiv N \frac{1-r_1 r_2}{1+r_1 r_2},
\end{equation}
\end{linenomath*}
where $N$ is the total number of samples (444 months for this timeseries), $r_1$ is the lag-1 autocorrelation for EKE and $r_2$ is the lag-2 autocorrelation for wind stress.
We use the Students t-test to infer statistical significance (at the 95\%-level) based on this effective sample size, when
\begin{linenomath*}
\begin{equation}
T = \frac{r \sqrt{N_e}}{\sqrt{1-r^2}} > 2.
\end{equation}
\end{linenomath*}
Lagged correlation estimates indicate when this significance test is satisfied.

@@

For each of the 50 members of the OCCIPUT ensemble we take the EKE averaged over the entire Southern Ocean (40$^\circ$-60$^\circ$S), and plot the deseasonalised EKE anomalies from each ensemble member as thin grey lines in Fig.~\ref{Fig:2}(a). 
This plot highlights the considerable spread in EKE, even when averaged over the full circumpolar belt; in other words, there is a significant component of chaotic (intrinsic) variability in the Southern Ocean eddy field.
Nonetheless, when averaged over all ensemble members (red line in Fig.~\ref{Fig:2}a) the existence of a coherent (forced) component of eddy variability is revealed.
Averaged over this wide region the fraction, $R_i$, of intrinsic variance is 0.82, confirming the visual impression that intrinsic processes dominate the variability in the eddy field averaged over the basin.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\hsize]{Figure2}
\caption{Eddy kinetic energy statistics over the Southern Ocean (40$^\circ$S--60$^\circ$S). (a) Spatially averaged deseasonalised EKE anomaly (relative to climatology) for individual ensemble members (grey) and the ensemble mean EKE anomaly (red) along with wind stress anomaly (black); and (b) Time-lagged correlation between wind stress and EKE for the ensemble mean (red) and individual ensemble  members (grey) -- with two individual ensemble members highlighted in magenta and orange.}
\label{Fig:2}
\end{center}
\end{figure}

Although Southern Ocean EKE variability is strongly intrinsic, there remains a component of forced variability (red line in Fig.~\ref{Fig:2}a).
Previous studies have inferred that there is a strong contribution of wind stress forcing upon EKE, and we therefore compare the forced variability with the variations in wind stress averaged over the same circumpolar belt (black line in Fig.~\ref{Fig:2}a).
This comparison suggests a relationship in which wind stress leads variations in EKE, consistent with previously published results \citep{Meredith-Hogg-2006,Hogg2015}.
However, the time-lagged correlations between wind stress and EKE implies that this relationship is complex. 
The intrinsically variable nature of Southern Ocean eddies means that for some ensemble members, there is no meaningful correlation between wind stress and eddies (grey lines in Fig~\ref{Fig:2}b).
On the other hand, ensemble member 5 (magenta line) has a clear (and significant; $T = 5.5$) correlation with a 4-month lag, while member 25 (orange line) is weakly correlated at 6 months, and significantly correlated ($T = 4.2$) at a 24-month lag.
These isolated examples highlight the differing behaviour of each ensemble member. 
The ensemble mean (red line in Fig~\ref{Fig:2}b) includes a significant correlation at $\sim$4 months and a second significant peak at $\sim$30 months.
Thus, two distinct timescales of response of the Southern Ocean eddy field to wind stress are present in this model, overprinted by a dominant and complex influence of chaotic variability.

>>>>

The surface eddy kinetic energy for each ensemble member is calculated as 
\begin{linenomath*}
\begin{equation}
E_i = \frac{1}{2}(u_i'^2 + v_i'^2),
\end{equation}
\end{linenomath*}
and subsampled at monthly temporal resolution through weighted averages.
It is notable that $E_i$ contains a large component of seasonal variation which dominates the statistics \citep{Martinez-Moreno2022}.
To look at the forced response, we therefore deseasonalise $E_i$, by removing the climatological mean; we eliminate noise on the monthly timescale by filtering with a 4-month rolling mean.
We then proceed to compute ensemble statistics from this deseasonalised EKE.

@@

To examine the relationship between the EKE (of either the ensemble mean, or from individual ensemble members) and its forcing, we are guided by previous work which emphasises the role of wind stress in governing the eddy variability at different timescales \citep[e.g.][]{Hogg2015, Sinha2016}.
Thus, we look at the correlation coefficient, $r$, at a range of lags between the wind stress $|\tau|$ (which is identical for all ensemble members and has been deseasonalised similarly to $E_i$) and EKE.
For each correlation we evaluate the statistical significance of the correlation \citep[following, e.g.][]{Santer2000}.
We first calculate the effective sample size, $N_e$, where
\begin{linenomath*}
\begin{equation}
N_e \equiv N \frac{1-r_1 r_2}{1+r_1 r_2},
\end{equation}
\end{linenomath*}
where $N$ is the total number of samples (444 months for this timeseries), $r_1$ is the lag-1 autocorrelation for EKE and $r_2$ is the lag-2 autocorrelation for wind stress.
We use the Students t-test to infer statistical significance (at the 95\%-level) based on this effective sample size, when
\begin{linenomath*}
\begin{equation}
T = \frac{r \sqrt{N_e}}{\sqrt{1-r^2}} > 2.
\end{equation}
\end{linenomath*}
Lagged correlation estimates indicate when this significance test is satisfied.

@@

For each of the 50 members of the OCCIPUT ensemble we take the EKE averaged over the entire Southern Ocean (40$^\circ$-60$^\circ$S), and plot the deseasonalised EKE anomalies from each ensemble member as thin grey lines in Fig.~\ref{Fig:2}(a). 
This plot highlights the considerable spread in EKE, even when averaged over the full circumpolar belt; in other words, there is a significant component of chaotic (intrinsic) variability in the Southern Ocean eddy field.
The observed EKE anomaly \citep[calculated using data from ][ and filtered using the same methods as model output]{Martinez-Moreno2022} indicates that observations are indistinguishable from the individual ensemble members.
Nonetheless, when averaged over all ensemble members (red line in Fig.~\ref{Fig:2}a) the existence of a coherent (forced) component of eddy variability is revealed.
Averaged over this wide region the fraction, $R_i$, of intrinsic variance is 0.82, confirming the visual impression that intrinsic processes dominate the variability in the eddy field averaged over the basin.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\hsize]{Figure2}
\caption{Eddy kinetic energy statistics over the Southern Ocean (40$^\circ$S--60$^\circ$S). (a) Spatially averaged deseasonalised EKE anomaly (relative to climatology) for individual ensemble members (grey) and the ensemble mean EKE anomaly (red) and observed EKE anomaly from the satellite era \citep[green line; calculated from ][]{Martinez-Moreno2022}, along with wind stress anomaly (black); and (b) Time-lagged correlation between wind stress and EKE for the ensemble mean (red) and individual ensemble  members (grey) -- with two individual ensemble members highlighted in magenta and orange.}
\label{Fig:2}
\end{center}
\end{figure}

Although Southern Ocean EKE variability is strongly intrinsic, there remains a component of forced variability (red line in Fig.~\ref{Fig:2}a).
Previous studies have inferred that there is a strong contribution of wind stress forcing upon EKE, and we therefore compare the forced variability with the variations in wind stress averaged over the same circumpolar belt (black line in Fig.~\ref{Fig:2}a).
This comparison suggests a relationship in which wind stress leads variations in EKE, consistent with previously published results \citep{Meredith-Hogg-2006,Hogg2015}.
However, the time-lagged correlations between wind stress and EKE implies that this relationship is complex. 
The intrinsically variable nature of Southern Ocean eddies means that for some ensemble members, there is no meaningful correlation between wind stress and eddies (grey lines in Fig~\ref{Fig:2}b).
On the other hand, ensemble member 5 (magenta line) has a clear (and significant; $T = 5.5$) correlation with a 4-month lag, while member 25 (orange line) is weakly correlated at 6 months, and significantly correlated ($T = 4.2$) at a 24-month lag.
These isolated examples highlight the differing behaviour of each ensemble member. 
The ensemble mean (red line in Fig~\ref{Fig:2}b) includes a significant correlation at $\sim$4~months ($T=5.9$) and a second significant peak at $\sim$30~months ($T=3.5$).
Thus, two distinct timescales of response of the Southern Ocean eddy field to wind stress are present in this model, overprinted by a dominant and complex influence of chaotic variability.
